---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README__",
  out.width = "90%",
  fig.align = "center",
  dev = "svg"
)
```

# unitquantreg: An R Package for Estimation and Inference on Parametric Quantile Regression Models for Bounded Data

<!-- badges: start -->
<!-- badges: end -->

The goal of `unitquantreg` is to provide tools for estimation and inference on
parametric quantile regression models for bounded data.

We developed routines with similar interface the `stats::glm` function,  which contains
estimation, inference, residual analysis, prediction, model comparison.

For more computation efficient the dpqr's, likelihoods, score and
hessian functions are vectorized and coded in `C++`.

The parameter estimation and inference are performed under the frequentist paradigm. 
The [**optimx**](https://cran.r-project.org/web/packages/optimx/index.html) R package
is use, since allows different optimization technique to maximize the log-likelihood function.
The analytical  score function are use in the maximization and the standard errors
are computed using the analytical  hessian matrix, both are implemented in efficient away using `C++`.



## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("AndrMenezes/unitquantreg")
```

## Example

The distribution families available are:

```{r}
lt_families <- list("unit-Weibull" = "uweibull",
                    "Kumaraswamy" = "kum",
                    "unit-Logistic" = "ulogistic",
                    "unit-Birnbaum-Saunders" = "ubs",
                    "log-extended Exponential-Geometric" = "leeg",
                    "unit-Chen" = "uchen",
                    "unit-Generalized Half-Normal-E" = "ughne",
                    "unit-Generalized Half-Normal-X" = "ughnx",
                    "unit-Gompertz" = "ugompertz",
                    "Johnson-SB" = "johnsonsb",
                    "unit-Burr-XII" = "uburrxii")
```


The workhorse function is `unitquantreg`, which follows same interface like `stats::glm`.

```{r example, }
library(unitquantreg)
data(water)
lt_fits <- lapply(lt_families, function(fam) {
  unitquantreg(
    formula = phpws ~ mhdi + incpc + region + log(pop), data = water, tau = 0.5,
    family = fam, link = "logit", link.theta = "log")
})
t(sapply(lt_fits, coef))
```

You can use the `likelihood_stats` to get likelihood-base statistics:

```{r}
likelihood_stats(lt = lt_fits)
```

It is also possible to perform pairwise [Voung test](https://en.wikipedia.org/wiki/Vuong%27s_closeness_test) to model selection of nonnested models.

```{r}
# Select just a few model to not mess the output
lt_chosen <- lt_fits[c("unit-Logistic", "Johnson-SB", "unit-Burr-XII", "unit-Weibull")]
pairwise.voung.test(lt = lt_chosen)
```


The currently methods implemented for `unitquantreg` objects are
```{r methods}
methods(class = "unitquantreg")
```


It is also possible to fit the model for several quantiles values like `rq` function from [**quantreg**](https://cran.r-project.org/web/packages/quantreg/index.html) package.

```{r various-quantiles}
fits <- unitquantreg(formula = phpws ~ mhdi + incpc + region + log(pop),
                     data = water, tau = 1:49/50, family = "uweibull",
                     link = "logit", link.theta = "log")
class(fits)
plot(fits, which = "coef", mean_effect = FALSE)
```

The currently methods implemented for `unitquantregs` objects are
```{r }
methods(class = "unitquantregs")
```
